---
title: "R4DS-Quickref: Ch9 -- Working with Strings: stringr"
output: html_notebook
---

```{r}
library(tidyverse)
```

### Counting String Patterns

The str_count() function counts how many tokens a string contain, where tokens, for 
example, can be characters or words. By default, str_count() will count the number of 
characters in a string.

```{r}
# Strings
(strings <- c(
 "Give me an ice cream",
 "Get yourself an ice cream",
 "We are all out of ice creams",
 "I scream, you scream, everybody loves ice cream.",
 "one ice cream,
 two ice creams,
 three ice creams",
 "I want an ice cream. Do you want an ice cream?"
))
```

```{r}
# Count individual strings (by default the count of all characters in the string)
str_count(strings)
```

```{r}
# So 'str_length' gives the same result here
str_length(strings)
```

```{r}
# But 'length' does not, of course
length(strings)
```

You can be explicit in specifying that str_count() should count characters by giving 
it a boundary() option. This determines the boundary between tokens (characters, 
words, lines, etc.), that is, the units to count.

```{r}
# Count strings, by character
str_count(strings, boundary("character"))
```

```{r}
# by words per string
str_count(strings, boundary("word"))
```

```{r}
# by sentences per line
str_count(strings, boundary("sentence"))
```

```{r}
# by number of lines
str_count(strings, boundary("line_break"))
```


```{r}
# Count substrings
str_count(strings, "ice cream")
```

```{r}
# Here you get 'scream' as well
str_count(strings, "cream")
```

The pattern you ask str_count() to count is not just a string. It is a regular 
expression. Some characters take on special meaning in regular expressions.For 
example, a dot represents any single character, not a full stop.

```{r}
# Count the number of any characters per line
str_count(strings, ".")
```

```{r}
# Count the literal string
str_count(strings, fixed("."))
```

```{r}
# Previous can be also done like so
str_count(strings, "\\.")
```

Since the pattern is a regular expression, we can use it to count punctuation 
characters:

```{r}
# Count punctuation characters
str_count(strings, "[[:punct:]]")
```

or the number of times ice cream(s) is at the end of the string:
```{r}
# The s? means zero or one s, and the $ means the end of the string.
str_count(strings, "ice creams?$")
```

Or rather, at the end of the string except that it might be followed by a 
punctuation mark.

```{r}
str_count(strings, "ice creams?[[:punct:]]?$")
```

### Splitting Strings

Sometimes, you want to split a string based on some separator—not unlike how we 
split on commas in comma-separated value files. The stringr function for this is str_
split()

```{r}
# New strings
(strings <- c(
 "one",
 "two",
 "one two",
 "one two",
 "one. two."
))
```

```{r}
# Split on space
# Note that this works even when there are more than one space
str_split(strings, " ")
```

You can use the boundary() function for splitting as well. For example, you can split 
a string into its words using boundary("word"):

```{r}
# We also get rid of the full stops in the last string.
str_split(strings, boundary("word"))
```

### Capitalizing Strings

You can use the str_to_lower() and friends to transform a string into different case.

```{r}
(macdonald <- "Old MACDONALD had a farm.")
```

```{r}
str_to_lower(macdonald)
str_to_upper(macdonald)
str_to_title(macdonald)
str_to_sentence(macdonald)
```

### Wrapping, Padding, and Trimming

If you want to wrap strings, that is, add newlines, so they fit into a certain width, you can 
use str_wrap().

```{r}
# New strings again
(strings <- c(
 "Give me an ice cream",
 "Get yourself an ice cream",
 "We are all out of ice creams",
 "I scream, you scream, everybody loves ice cream.",
 "one ice cream,
 two ice creams,
 three ice creams",
 "I want an ice cream. Do you want an ice cream?"))
```

```{r}
str_wrap(strings)
```

```{r}
# The default width is 80 characters, but you can change that using the width argument.
str_wrap(strings, width = 10) 
```

```{r}
# You can indent the first line in the strings while wrapping them using the indent argument.
str_wrap(strings, width = 10, indent = 2)
```

If you want your string to be left, right, or center justified, you can use str_pad().
The default is right-justifying strings.

```{r}
# Default is right-justify
str_pad(strings, width = 50)
```

```{r}
# Right-justify actually uses left-padding
str_pad(strings, width = 50, side = "left")
```

```{r}
# If you want to left-justify instead, you can pass "right" to the side argument.
str_pad(strings, width = 50, side = "right")
```

You need to use "right" to left-justify because the side argument determines which 
side to pad, and for left-justified text, the padding is on the right.
If you want to center your text, you should use "both"; you are padding both on the 
left and on the right.

```{r}
# Center text
str_pad(strings, side = "both", width = 50)
```

In these padding examples, we do not keep the lengths of the strings below the 
padding width. If a string is longer than the padding width, it is unchanged. You can 
use the str_trunc() function to cut the width down to a certain value. For example, we 
could truncate all the strings to width 25 before we pad them:

```{r}
strings %>% 
  str_trunc(25) %>% 
  str_pad(width = 25, side = "left")
```

The str_trim() function removes whitespace to the left and right of a string:

```{r}
# Cokes
(cokes <- c(
  " one small coke",
  "two large cokes ",
  " three medium cokes "
))
```

```{r}
# Whitespace inside the string is preserved.
str_trim(cokes, side = "both")
```

For example, if we have two  spaces between two words, 
as follows, str_trim() leaves them alone:

```{r}
(cokes <- c(
  " one small coke",
  "  two   large cokes ",
  " three medium cokes   "
))
```

```{r}
str_trim(cokes, side = "both")
```

If we want the two spaces to be shortened into a single space, we can use str_
squish() instead.

```{r}
# This also trims the spaces at the ends
str_squish(cokes)
```

### Detecting Substrings

To check if a substring is found in another string, you can use str_detect().

```{r}
str_detect(strings, "me")
str_detect(strings, "I")
str_detect(strings, "cream")
```

```{r}
# You can do an inverse search using 'negate = TRUE'
str_detect(strings, "I")
str_detect(strings, "I", negate = TRUE)
```

The pattern is a regular expression, so to test for ice cream followed by a full stop, you 
cannot search for “ice cream.”

```{r}
# You can use 'fixed' or escaping to get the correct result
str_detect(strings, "ice cream.")
str_detect(strings, fixed("ice cream."))
str_detect(strings, "ice cream\\.")
```

Two special case functions test for a string at the start or end of a string:

```{r}
str_starts(strings, "I")
str_ends(strings , "\\.") # escaping to search literal dot
```

If you want to know where a substring is found, you can use str_locate(). It will 
give you the start and end index where it found a match.

```{r}
# Here, you get a start and end index for the first occurrence of each string 
str_locate(strings, "ice cream")
```

The function str_locate_all() gives you all occurrences.
```{r}
str_locate_all(strings, "ice cream")
```

If you want the start and end points of the strings between the occurrences, you can 
use invert_match().

```{r}
# Using map to get the inverted matches for all
all_matches <- str_locate_all(strings, "ice cream")  
map(all_matches, invert_match)
```

### Extracting Substrings

To extract a substring matching a pattern, you can use str_extract(). It gives you the 
first substring that matches a regular expression.

```{r}
# Match the first "scream" or "ice cream" and 0 or more word characters
str_extract(strings, "(s|ice )cream\\w*")
```

```{r}
# Get all matches
str_extract_all(strings, "(s|ice )cream\\w*")
```

### Transforming Strings

We can replace a substring that matches a pattern with some other string.

```{r}
(lego_str <- str_replace(strings, "ice creams?", "LEGO"))
```

```{r}
lego_str %>% str_replace("an LEGO", "a LEGO")
```

These two replacement operators can be written as a pipeline to make the code more
```{r}
strings %>% 
  str_replace("ice creams?", "LEGO") %>% 
  str_replace("an LEGO", "a LEGO")
```

```{r}
# To replace all occurrences, you need str_replace_all().
strings %>% 
  str_replace_all("ice creams?", "LEGO") %>% 
  str_replace_all("an LEGO", "a LEGO")
```

You can refer back to matching groups in the replacement string, something you will 
be familiar with for regular expressions.

```{r}
(us_dates <- c(
 valentines = "2/14",
 my_birthday = "2/15",
 # no one knows but let's just go with this
 jesus_birthday = "12/24"
))
```

```{r}
# US date format to a more sane format
# Capture groups the reorder them (first = \\1, second = \\2)
str_replace(us_dates, "(.*)/(.*)", "\\2/\\1")
```

The str_dup() function duplicates a string, that is, it repeats a string multiple times.

```{r}
# Here, we used str_c() to concatenate strings. This is a stringr version of paste.
str_c("NA", str_dup("-NA", times = 7), " BATMAN!")
```

A more direct way to extract and modify a substring is using str_sub(). It lets 
you extract a substring specified by a start and an end index, and if you assign to it, 
you replace the substring. The str_sub() function is less powerful than the other 
functions as it doesn’t work on regular expressions, but because of this, it is also easier to 
understand.

```{r}
# Use str_locate to get the start and end indexes
my_string <- "this is my string"
my_location <- str_locate(my_string, "my")
my_location
```

```{r}
# Use str_sub to extract substring
s <- my_location[,"start"]
e <- my_location[,"end"]
str_sub(my_string, s, e)
```

```{r}
my_string_location <- str_locate(my_string, "string")
s <- my_string_location[,"start"]
e <- my_string_location[,"end"]
str_sub(my_string, s, e)
```

```{r}
# How the hell does this work?
(your_string <- my_string)
(s <- my_location[,"start"])
(e <- my_location[,"end"])
(str_sub(your_string, s, e) <- "your") # This part??
your_string
```

```{r}
# How the hell does this work?
(your_banana <- your_string)
(your_string_location <- str_locate(your_string, "string"))
(s <- your_string_location[,"start"])
(e <- your_string_location[,"end"])
(str_sub(your_banana, s, e) <- "banana") # This part??
your_banana
```

When you assign to a call to str_sub(), it looks like you are modifying a string. This 
is an illusion. Assignment functions create new data and change the data that a variable 
refers to. So, if you have more than one reference to a string, be careful. Only one variable 
will point to the new value; the remaining will point to the old string. This is not specific 
to str_sub() but for R in general, and it is a potential source of errors.

If you often write code to produce standard reports, virtually the same text each time 
but with a few selected values changed, then you are going to love str_glue(). This 
does precisely what you want. You give str_glue() a template string, a string with the 
mutable pieces in curly brackets. The result of calling str_glue() is the template text but 
with the values in the curly brackets replaced by what R expression they contain.
The most straightforward use is when the template refers to variables.

```{r}
macdonald <- "Old MacDonald"
eieio <- "E-I-E-I-O"
```

```{r}
# This is like those f-strings in python
str_glue("{macdonald} had a farm. {eieio}")
```

```{r}
# The variables do not need to be global. T
str_glue(
 "{macdonald} had a farm. {eieio}",
 macdonald = "Thomas",
 eieio = "He certainly did not!"
)
```

Generally, you can put R expressions in the curly brackets, and the result of 
evaluating the expressions will be what is inserted into the template string.

```{r}
str_glue("{str_dup(\"NA-\", times = 7)}NA BATMAN!")

str_glue("Holy {mean(1:10)} BATMAN!")
```

### Regular expressions

```{r}
my_words <- c( "cat", "cart","carrot", "catastrophe",
               "dog","rat",  "bet")
my_words
```

```{r}
# Flexible matching though metacharacters
str_view(my_words , ".at")
```

```{r}
# . on it's own matches exactly one occurrence
str_subset(my_words, "c..t")
```

```{r}
# + (plus) represents one or more occurrences
str_subset(my_words, "c.r+")
```

```{r}
# * (star) represents zero or more occurrences
str_subset(my_words, "c.r*")
```

```{r}
# Group terms with parentheses ( and )
str_extract(my_words, "(at)|(og)+")
```

```{r}
# Alternation operator | ( logical OR ) for groups
str_match(my_words, "(c.t)|(c.rt)")
```

```{r}
# Capture groups with str_match()
str_match(my_words, "c(a.)+t")
```

```{r}
# Quantifying a number of matches
# A or AA
dna <- "ATGGTAACCGGTAGGTAGTAAAGGTCCC"
str_view(dna, "AA?")
```

```{r}
# AA or more A's
str_view(dna, "AA+")
```

```{r}
# Three or more A's
str_view(dna, "A{3,}")
```

```{r}
# Matches are greedy by default
# Match the longest possible subsequence.
str_extract(dna, "AAG.{2,5}")
str_extract(dna, "ATG.+C")
```

```{r}
# Lazy matching with ?
str_extract(dna, "AAG.{2,5}?")
str_extract(dna, "ATG.+?C")
```

```{r}
# Anchors
# ^ Start of string
my_words
str_subset(my_words , '^c')
```

```{r}
# $ End of string
str_subset(my_words, "r.$")
```

Character classes

```{r}
# Examples

uniprot <- c("Q6QU88_CALBL", "CO1A2_HUMAN", "SAMH1_HUMAN", 
             "NPRL2_DROME", "GLUC_HUMAN")

str_extract(uniprot, "\\d+")
```

```{r}
str_count(uniprot, "\\w+")
```

```{r}
# Matching "everything that you want"
str_extract(uniprot, "\\S+_")
```

```{r}
str_subset(uniprot, "[:digit:]")
str_extract(uniprot, "[:digit:]")
```

```{r}
# Matching metacharacters
# Use \\ to escape to literal character (in many other langs only \ is used)
vec2 <- c("YKL045W-A", "12+45=57", "$1200.00", "ID2.2")

# Not what we want
str_subset(vec2 , '.')

# Better
str_subset(vec2 , '\\.')
```

```{r}
# Backreferences
# Group matches: \1, \2 and so forth refer to groups matched with ().
str_replace(uniprot, '(\\S+)_(\\S+)', "\\2: \\1")
```

Helpers
regexplain

Simple addin for RStudio by Garrick Aden-Buie

    Test regular expressions on the fly
    Reference library
    Cheatsheet
    test it live at https://www.garrickadenbuie.com/project/regexplain/


devtools::install_github("gadenbuie/regexplain")
regexplain::regexplain_gadget()