---
title: "R4DS-Quickref: Ch7 -- Functional Programming: purrr"
output: html_notebook
---

```{r}
library(tidyverse)
library(magrittr)
```


### Filtering

One of the most straightforward functional programming patterns is filtering. Here, 
you use a high-order filter function. This function takes a predicate function, that is, a 
function that returns a logical value, and then returns all elements where the predicate 
evaluates to TRUE or all elements where the predicate evaluates to FALSE.

```{r}
# Keep even numbers only
is_even <- \(x) x %% 2 == 0

1:10 %>% keep(is_even)
```

```{r}
# Keep odd numbers only

1:10 %>% discard(is_even)
```

When you work with predicates, the negate() function can come in handy. It 
changes the truth value of your predicate, such that if your predicate, p, returns TRUE, 
then negate(p) returns FALSE and vice versa.

```{r}
# Keep even
1:10 %>% discard(negate(is_even))
```

```{r}
# Keep odd
1:10 %>% keep(negate(is_even))
```

Since you already have complementary functions in keep() and discard(), you 
would not use negate() for filtering, though.


`compact()` removes NULL elements from a list

```{r}
# List with empty elements
(y <- list(NULL, 1:3, NULL))
```

```{r}
# Remove NULLs
compact(y)
```

If you access attributes on objects, and those are not set, R will give you NULL as a 
result, and for such cases, compact() can be useful.

```{r}
# Define two vectors
x <- y <- 1:3
x
y
```


```{r}
# Name the elements of one vector, but keep the other unnamed
names(y) <- c("one", "two", "three")
names(y)
names(x)
```

```{r}
# Create a list
(z <- list(x = x, y = y))
```

```{r}
# Remove vector with unnamed elements
z %>% compact(names)
```

### Mapping

Mapping a function, f, over sequences x = x1, x2, …, xn returns a new sequence of the 
same length as the input but where each element is an application of the function: f(x1), 
f(x2), …, f(xn).

The function map() does this and returns a list as output. Lists are the generic 
sequence data structure in R since they can hold all types of R objects.

```{r}
# Compare the output of using plain is_even() on the vector and then using map().
# Mapping returns a list
1:4 %>% is_even()
1:4 %>% map(is_even)
```

Often, we want to work with vectors of specific types. For all the atomic types, purrr
has a specific mapping function. The function for logical values is named map_lgl().

```{r}
# map_lgl() returns a logical vector, like plain is_even()
1:4 %>% is_even()
1:4 %>% map_lgl(is_even)
```

With something as simple as this example, you should not use purrr. Vector 
expressions are faster and easier to use.

```{r}
1:4 %% 2 == 0
```


You cannot always use vector expressions, however. Say you have a list of vectors like below:

```{r}
# A list of vectors
(my_list <- list(1:5, 200:540, 5:60))
```

```{r}
# This does not give the lengths of individual vectors in the list
my_list %>% length()
```

When you have a function that is not vectorized, you can use a map function to apply 
it on all elements in a list.

```{r}
# This works
my_list %>% map(length)
```

```{r}
# If you want it as a integer vector, you can use map_int()
my_list %>% map_int(length)
```

There are many versions of `map_*()`

map(.x, .f, ..., .progress = FALSE)

map_lgl(.x, .f, ..., .progress = FALSE)

map_int(.x, .f, ..., .progress = FALSE)

map_dbl(.x, .f, ..., .progress = FALSE)

map_chr(.x, .f, ..., .progress = FALSE)

map_vec(.x, .f, ..., .ptype = NULL, .progress = FALSE) !This is new in purrr 1.0.0

walk(.x, .f, ..., .progress = FALSE) !Side effects

```{r}
# identity() is trivial identity function returning its argument. 
1:3 %>% map_dbl(identity) %T>% print() %>% class() 
```

```{r}
1:3 %>% map_int(identity) %T>% print() %>% class() 
```

```{r}
1:3 %>% map_chr(identity) %T>% print() %>% class() 
```

```{r}
# map_vec() is a general mapper for eg. dates, factors etc.
1:3 %>% map_vec(\(x) Sys.Date() + x)
letters[1:3] %>% map_vec(\(x) factor(x))
```


The different map functions will give you an error if the function you apply does not 
return values of the type the function should. Sometimes, a map function will do type 
conversion, as before, but not always. It will usually be happy to convert in a direction 
that doesn’t lose information, for example, from logical to integer and integer to double, 
but not in the other direction.

!! NOTE THE `map_df*()` functions below have been deprecated. 
See further below how to do this now !!
The `map_dfr()` and `map_dfc()` functions return data frames (tibbles). 
The `map_df()` function does the same as `map_dfr()`.

```{r}
# map_dfr and map_df bind the output by rows
x <- tibble(a = 1:2, b = 3:4)
list(a = x, b = x) %>%  map_dfr(identity)
```

```{r}
# map_df also binds by rows
list(a = x, b = x) %>%  map_df(identity)
```

```{r}
# map_dfc binds the output by columns
list(a = x, b = x) %>%  map_dfc(identity)
```

The suggested method to do the above procedures is using `map()` and `list_rbind()` or `list_cbind()`

```{r}
# Bind output data frames by rows
x <- tibble(a = 1:2, b = 3:4)
list(a = x, b = x) %>%  
  list_rbind()
```

```{r}
# dplyr's bind_rows() also seems to give the same output
list(a = x, b = x) %>% 
  bind_rows()
```


```{r}
# list_cbind creates list columns in this case
list(a = x, b = x) %>%  
  map(identity) %>% 
  list_cbind()
```

```{r}
# With bind_cols() you get a similar result as with mad_dfc()
list(a = x, b = x) %>%  
  bind_cols() 
```

You do not need to give the data frame functions data frames as input, as long as the 
function you apply to the input returns data frames. This goes for all the map functions. 
They will accept any sequence input; they only restrict and convert the output.

If the items you map over are sequences themselves, you can extract elements by 
index; you do not need to provide a function to the map function.

The three below give the same result:
```{r}
# Extract by index only using map()
list(1:3, 4:6) %>% 
   map_dbl(1) # First element of each list

list(1:3, 4:6) %>% 
   map_dbl(3) # Third element of each list
```

pluck() implements a generalised form of [[ that allow you to index deeply and flexibly into data structures. It always succeeds, returning .default if the index you are trying to access does not exist or is NULL.

```{r}
# Same as x[[1]][[2]]
list(1:3, 4:6) %>% 
  pluck(1, 2) # Second element of the first element 
```

```{r}
# Non-existing element returns NULL
list(1:3, 4:6) %>% 
  pluck(1, 6)
```

```{r}
# Use pluck() together with map()
# Pluck is mostly used for digging deeper into nested lists
list(1:3, 4:6) %>%
   map_dbl(pluck(1))
```

```{r}
# Map `[` or `[[` to extract element
list(1:3, 4:6) %>% 
  map_dbl(`[[`, 1 )
```

If the items have names, you can also extract values using these.
```{r}
# Use names to extract
(x <- list(
  c(a = 42, b = 13),
  c(a = 24, b = 31)
))

x %>% map_dbl("a")
x %>% map_dbl(pluck("b"))
x %>% map_dbl(`[[`, "a")
```

This is mostly used when you map over data frames.
```{r}
a <- tibble(foo = 1:3, bar = 11:13)
b <- tibble(foo = 4:6, bar = 14:16)
(ab <- list(a = a, b = b))
```

```{r}
ab %>% map("foo")
```

You can use `pluck()` to extract from deeper levels
```{r}
# Make a combo list for demo
my_list <- list(a = iris, b = mtcars)
```

```{r}
# First is just the top level list elements
my_list %>% pluck(1)
my_list %>% pluck(2)
```

```{r}
# Second level is the columns in the individual element data.frames
my_list %>% pluck(1, 2)
my_list %>% pluck(2, 2)
```

```{r}
# Third level is the individual elements in the data frame vectors
my_list %>% pluck(1, 2, 5)
my_list %>% pluck(2, 2, 5)
```

```{r}
# Using map_dbl to extract the fifth element of the second column in both data frames
my_list %>% 
  map_dbl(\(x) pluck(x, 2, 5))
```

Related to extracting elements with pluck, you can apply functions to different 
depths of the input using map_depth(). As with pluck, depth zero is the list itself, 
so mapping over this depth is the same as applying the function directly on the input.

Depth 1 gives us each element in the sequence, so this behaves like a normal map. 
Depth 2 provides us with a map over the nested elements. Consider `my_list` (a = iris, b = mtcars). 
The top level, depth 0, is the list. Depth 1 is the data frames `iris` and `mtcars`. 
Depth 2 is the columns in these data frames. Depth 3 is the individual items in these columns.

    map_depth(x, 0, fun) is equivalent to fun(x).

    map_depth(x, 1, fun) is equivalent to x <- map(x, fun)

    map_depth(x, 2, fun) is equivalent to x <- map(x, \(y) map(y, fun))


```{r}
# map_depth(x, 0, fun) is equivalent to fun(x).
map_depth(my_list, 0, names)
names(my_list)

map_depth(my_list, 0, length)
length(my_list)
```

```{r}
# map_depth(x, 1, fun) is equivalent to map(x, fun)
map_depth(my_list, 1, names)
map(my_list, names)

map_depth(my_list, 1, length)
map(my_list, length)
```

```{r}
# NOTE! map_depth(x, 2, fun) requires .ragged = TRUE to work here, why?
# map_depth(x, 2, fun) is equivalent to map(x, \(y) map(y, fun))
# map_depth(my_list, 2, identity)

map(my_list, \(x) map(x, length))
map_depth(my_list, 2, length, .ragged = TRUE)
```

```{r}
# Depth 0 is just like using the function normally.
ab %>% length()
ab %>% map_depth(0, length)
```

```{r}
# Depth 1 is like normal map.
ab %>% map_depth(1, sum)
```

```{r}
# Depth 2 is for the individual columns in the dataframes in this case
ab %>% map_depth(2, sum, .ragged = TRUE) # Why the .ragged = TRUE is needed?
```

```{r}
# Depth 3 is the individual elements in the dataframe columns
ab %>% map_depth(3, \(x) x %% 2 == 0, .ragged = TRUE) 
```

If you only want to apply a function to some of the elements, you can use `map_if()`. 
It takes a predicate and a function and applies the function to those elements where the 
predicate is true. It returns a list, but you can convert it if you want another type.

```{r}
is_even <- \(x) x %% 2 == 0
add_one <- \(x) x + 1
```

```{r}
# Map function based on predicate. Returns a list. 
map_if(1:10, is_even, add_one)
```
With map_if(), you keep all elements, but the function is only applied to some 
of them.
```{r}
# Return a numeric vector with extra step
map_if(1:10, is_even, add_one) %>% as.numeric()
```

Notice that this is different from combining filtering and mapping; that combination 
would remove the elements that do not satisfy the predicate.
```{r}
# Filter and map
1:6 %>%  keep(is_even) %>% map_dbl(add_one)
```

Using modify_if() instead of map_if() automatically returns a similar object with modification

```{r}
# Return a numeric vector using modify_if()
modify_if(1:10, is_even, add_one)
```


With `map_if()` and `modify_if()`, you keep all elements, but the function is only applied to some 
of them. If you want to apply one function to the elements where the predicate is 
true and another to the elements where it is false, you can prove a function to the `.else` element:

```{r}
# Use .else = TRUE, to apply another function on elements that return FALSE
add_two <- \(x) x + 2

1:6  %>% 
 map_if(is_even, add_one, .else = add_two)  %>% 
 as.numeric()

# With modify_if
1:6 %>% modify_if(is_even, add_one, .else = add_two)
```

If you know which indices you want to apply the function to, instead of a predicate 
they must satisfy, you can use `map_at()`. This function takes a sequence of indices 
instead of the predicate but otherwise works the same as `map_if()`.

```{r}
# map_at and modify_at
1:6 %>%  map_at(2:5, add_one) %>%  as.numeric()
1:6 %>%  modify_at(2:5, add_one)
```

In purrr 1.0.0 you can use keep_at or discard_at to keep/discard elements by name
```{r}
1:6 %>%  discard_at(2:5)
1:6 %>%  keep_at(2:5)
```

```{r}
# Discard or keep based on element names
(myls <- list(a = 1, b = 4, c = 7, d = 9))

myls %>% discard_at(c("b", "d"))
```


If you map over a list, x, then your function will be called with the elements in the 
list, x[[i]]. If you want to get the elements wrapped in a length-one list, that is, use 
indexing x[i], you can use lmap().

```{r}
list(a = 1:3, b = 4:6) %>% map(print) %>% invisible()
```

```{r}
list(a = 1:3, b = 4:6) %>% lmap(print) %>% invisible()
```

The function you apply must always return a list, and lmap() will concatenate them.

```{r}
# This just returns the string "foo" inside a list
f <- function(x) list("foo")
```

```{r}
1:2 %>% lmap(f)
```

```{r}
f <- function(x) list("foo", "bar")
1:2 %>% lmap(f)
```

For example, while you can get the length of elements in a list using map() and 
length()

```{r}
# map and unlist
list(a = 1:3, b = 4:8) %>% map(length) %>% unlist()
```

```{r}
# map_int
list(a = 1:3, b = 4:8) %>% map_int(length)
```

you will get an error if you try the same using lmap(). This is because length()
returns a numeric and not a list. You need to wrap length() with list() so the result is 
the length in a (length one) list.

```{r}
list(a = 1:3, b = 4:8) %>% lmap(\(x) length(x) %>% list())
```

If it surprises you that the lengths are one here, remember that the function is called 
with the length-one lists at each index. If you want the length of what they contain, you 
need to extract that.

```{r}
list(a = 1:3, b = 4:8) %>% lmap(\(x) pluck(x, 1) %>% length() %>% list())
```

If you want to extract the nested data, though, you probably want map() and not lmap().


Sometimes, we only want to call a function for its side effect. In that case, you can 
pipe the result of a map into invisible(). The function walk() does that for you, and 
using it makes it explicit that this is what you want, but it is simply syntactic sugar for 
map() + invisible().

```{r}
1:3 %>% map(print) %>% invisible()
```

```{r}
1:3 %>% walk(print)
```

If you need to map over multiple sequences, you have two choices of map functions 
to choose from. Some functions map over exactly two sequences. For each of the map()
functions, there are similar map2() functions. These take two sequences as the first two 
arguments.

```{r}
# map2
x <- 1:3
y <- 3:1
map2_dbl(x, y, `+`)
x %>% map2_dbl(y, `+`)
```

```{r}
# pmap
pmap_dbl(list(x, y), `+`)
list(x, y) %>% pmap_dbl(`+`)
```

```{r}
# With pmap you can use more than two sequences
z <- 4:6

list(x, y, z) %>% pmap_dbl(\(x, y, z) x + y - z)
```

If you need to know the indices for each value you map over, you can use the imap()
variations. When you use these to map over a sequence, your function needs to take two 
arguments where the first argument is the sequence value and the second the value’s 
index in the input.

```{r}
x <- c("foo", "bar", "baz")

x %>% imap_chr(\(x, i) paste0(i, ": ", x))
```

There is yet another variant of the mapping functions, the modify() functions. These 
do not have the type variants (but the _at, _if, _depth, and so on); instead, they will 
always give you an output of the same type as the input:

```{r}
modify2(1:3, 3:1, `+`)
map2_dbl(1:3, 3:1, `+`)
```

```{r}
x <- c("foo", "bar", "baz")

x %>% imodify(\(x, i) paste0(i, ": ", x))
x %>% imap_chr(\(x, i) paste0(i, ": ", x))
```

### Reduce and Accumulate

If you want to summarize all your input into a single value, you probably want to 
reduce() them. Reduce repeatedly applies a function over your input sequence. If you 
have a function of two arguments, f (a, x), and a sequence x1, x2, …, xn, then reduce(f)
will compute f(… f( f(x1, x2), x3), …, xn), that is, it will be called on the first two elements 
of the sequence, the result will be paired with the next element, and so forth. Think of the 
argument a as an accumulator that keeps the result of the calculation so far.

```{r}
# Reduce and Accumulate
1:5 %>% reduce(`+`)
1:5 %>% accumulate(`+`)

letters[1:5] %>% reduce(paste, sep = ".")
letters[1:5] %>% accumulate(paste, sep = ".")
```


To make the order of function application clear, I define a “pair” type:
```{r}
pair <- function(first, second) {
 structure(list(first = first, second = second),
 class = "pair")
}

toString.pair <- function(x, ...) {
 first <- toString(x$first, ...)
 rest <- toString(x$second, ...)
 paste('[', first, ', ', rest, ']', sep = '')
}

print.pair <- function(x, ...) {
 x %>% toString() %>% cat() %>% invisible()
}
```

If we reduce using pair(), we see how the values are paired when the function 
is called:

```{r}
1:4 %>% reduce(pair)
```

If you reverse the input, you can reduce in the opposite order, combining the last pair 
first and propagating the accumulator in that order

```{r}
1:4 %>% rev() %>% reduce(pair)
```

If, for some reason, you want to apply the function and have the accumulator as the 
last argument, you can use the .dir = “backward” argument.

```{r}
1:4 %>% reduce(pair, .dir = "backward")
```

The first (or last) element in the input does not have to be the value for the initial 
accumulator. If you want a specific starting value, you can pass that to reduce() using 
the .init argument.

```{r}
1:3 %>% reduce(pair, .init = 0)
```

```{r}
1:3 %>% rev() %>% reduce(pair, .init = 4)
```

```{r}
1:3 %>% reduce(pair, .init = 4, .dir = "backward")
```

If your function takes more than one argument, you can provide the additional 
arguments to reduce() and then input sequence and function. Consider, for example, a 
three-argument function like this:
```{r}
# additional arguments
loud_pair <- function(acc, next_val, volume) {
 # Build a pair
 ret <- pair(acc, next_val)
 # Announce that pair to the world
 ret %>% toString() %>% 
 paste(volume, '\n', sep = '') %>% 
 cat()
 # Then return the new pair
 ret
}
```

It builds a pair object but, as a side effect, prints the pair followed by a string that 
indicates how “loud” the printed value is. We can provide the volume as an extra 
argument to reduce():

```{r}
1:3 %>% 
  reduce(loud_pair, volume = '!') %>% 
  invisible()
```

```{r}
1:3 %>% 
  reduce(loud_pair, volume = '!!') %>% 
  invisible()
```

If you want to reduce two sequences instead of one — similar to a second argument to 
reduce() but a sequence instead of a single value — you can use reduce2():

```{r}
# Second argument is one shorter than the other
volumes <- c('!', '!!')
1:3 %>% reduce2(volumes, loud_pair) %>% invisible()
```

```{r}
# Unless you have a specific initiator value
volumes <- c('!', '!!', '!!!')
1:3 %>% reduce2(volumes, loud_pair, .init = 0) %>% invisible()
```

```{r}
1:3 %>% 
  reduce2(c('!', '!!', '!!!'), .init = 0, loud_pair) %>% 
  invisible()
```

```{r}
reduce2(1:10, letters[1:9], ~ paste0(..1, ..2, ..3))
```

If you want all the intermediate values of the reductions, you can use the 
accumulate() function. It returns a sequence of the results of each function application.

```{r}
# Accumulate
1:5 %>% accumulate(pair)
```

```{r}
1:5 %>% accumulate(pair, .init = 0)
```

```{r}
1:4 %>% accumulate(pair, .init = 0, .dir = "backward")
```

The accumulate2() function works like reduce2(), except that it keeps the 
intermediate values like accumulate() does.

```{r}
accumulate2(1:10, letters[1:9], ~ paste0(..1, ..2, ..3))
```

```{r}
accumulate2(1:10, letters[1:10], ~ paste0(..1, ..2, ..3), .init = 0)
```


### Partial Evaluation and Function Composition

When you `filter`, `map`, or `reduce` over sequences, you sometimes want to modify a 
function to match the interface of `purrr`’s functions. If you have a function that takes 
too many arguments for the interface, but where you can fix some of the parameters to 
get the application you want, you can do what is called a partial evaluation. This just 
means that you create a new function that calls the original function with some of the 
parameters fixed.

For example, if you `filter`, you want a function that takes one input value and returns 
one (Boolean) output value. If you want to filter the values that are less than or greater 
than, say, three, you can create functions for this.

```{r}
greater_than_three <- \(x) x > 3
less_than_three <- \(x) x < 3
```

```{r}
1:6 %>% keep(greater_than_three)
1:6 %>% keep(less_than_three)
```

The drawback of doing this is that you might need to define many such functions, 
even if you only use each once in your pipeline.

Using the `partial()` function, you can bind parameters without explicitly defining 
new functions. For example, to bind the first parameter to `<`, as in the `greater_than_
three()` function, you can use `partial()`:

```{r}
partial(`<`, 3)
```


```{r}
# Greater than 3 with partial
1:6 %>% keep(partial(`<`, 3))
```

By default, you always bind the first parameter(s). To bind others, you need to name 
which parameters to bind. The less than operator has these parameter names:

```{r}
`<`
```

so you can use this partial evaluation for less_than_three():

```{r}
# Less than three with partial
1:6 %>% keep(partial(`<`, e2 = 3))
```

You can use partial evaluation for mapping:

```{r}
1:6 %>% map_dbl(partial(`+`, 2)) # 2 + 1, 2 + 2, 2 + 3, 2 + 4, 2 + 5, 2 + 6 
1:6 %>% map_dbl(partial(`-`, 1))
1:6 %>% map_dbl(partial(`-`, e2 = 4)) # 1 - 4, 2 - 4, 3 - 4, 4 - 4, 5 - 4, 6 - 4
```

```{r}
# More than one function
1:3 %>% 
  map_dbl(partial(`+`, 2)) %>% 
  map_dbl(partial(`*`, 3))
```

Using function composition with `compose()`, the previous can be written as:
```{r}
1:3 %>% map_dbl(compose(partial(`*`, 3), partial(`+`, 2)))
```

With `partial()` and `combine()`, you can modify functions, but using them does not 
exactly give you code that is easy to read. A more readable alternative is using lambda 
expressions.

### Lambda Expressions

Lambda expressions are a concise syntax for defining anonymous functions, that is, 
functions that we do not name. The name “lambda expressions” comes from “lambda 
calculus", a discipline in formal logic, but in computer science, it is mostly used as a 
synonym for anonymous functions.

In R, you define anonymous functions 
the same way that you define named functions. You always define functions the same 
way; you only give them a name when you assign a function definition to a variable.

In R v.4 you can use `\(x)` as a shortcut for `function(x)`

We can use a function without assigning it to a name several ways in R:
```{r}
# These are the same
1:6 %>% keep(less_than_three) # Fn is assigned to name
1:6 %>% keep(\(x) x < 3)
1:6 %>% keep(function(x) x < 3)
1:6 %>% keep(~ .x < 3) # purrr style formula
```

Normal lambda's work on their own (they are identical to function()), but purrr's
lambdas do not work outside a purrr function (or certain other Tidyverse functions).

```{r}
# This works
is_even_lambda <- \(x) x %% 2 == 0
is_even_lambda(3)
```

```{r}
# This does not work, as no function is defined
is_even_purrr <- ~ .x %% 2 == 0
is_even_purrr(3)
```

R’s rule for looking for functions can be even more confusing if there is a variable in 
an inner scope and a function in an outer scope:

```{r}
f <- function(x) 2 * x

g <- function() {
 f <- 5 # not a function
 f(2) # will look for a function, and finds it outside the scope of the function
}
g()
```

Here, the f() function in the outer scope is called because it is a function; the 
variable in the inner scope is ignored.

```{r}
g <- function() {
 f <- 5 # not a function
 f(f) # function f and variable f are not the same
}
g()
```

### purrr's as_mapper

https://purrr.tidyverse.org/reference/as_mapper.html

as_mapper is the powerhouse behind the varied function specifications that most purrr functions allow.

```{r}
# as_mapper creates a function from a formula (happens automatically when used with purrr?)
as_mapper(~ .x %% 2 == 0)
```

```{r}
# Now we can use the formula syntax fuction normally
is_even_purrr <- as_mapper(~ .x %% 2 == 0)
as_mapper(~ .x %% 2 == 0) (3) 
as_mapper(~ .x %% 2 == 0) (4)
is_even_purrr(5)
is_even_purrr(6)
```

If the expression inside as_mapper is a function, it is used as is.
If a formula, e.g. ~ .x + 2, it is converted to a function. There are three ways 
to refer to the arguments:

    For a single argument function, use .

    For a two argument function, use .x and .y

    For more arguments, use ..1, ..2, ..3 etc

This syntax allows you to create very compact anonymous functions. Note that 
formula functions conceptually take dots (that's why you can use ..1 etc). 
They silently ignore additional arguments that are not used in the formula expression.

```{r}
# Formula is made into a function
as_mapper(~ .x < 3)
```

```{r}
# A function is used as it is
as_mapper(\(x) x < 3)
```

```{r}
# A vector or list is made into an extractor function (note the use of pluck)
as_mapper(c(2, 3))
as_mapper(1)
as_mapper(c("a", "b", "c"))
as_mapper(list(1, "a", 2))
```

Behind the scenes `as_mapper` uses `rlang::as:function`

```{r}
rlang::as_function(~ .x < 3)
```

Examples of using the purrr style for readability

```{r}
1:4 %>% map_dbl(~ .x / 2)
1:3 %>% map_dbl(~ 2 + .x)
1:3 %>% map_dbl(~ .x - 4)
```

```{r}
# Or you can use them for more readable versions of function composition.
1:3 %>% map_dbl(~ 3 * (.x + 2))
```

```{r}
# If you need a lambda expression with two arguments, you can use .x and .y as the 
# first and second arguments, respectively.
map2_dbl(1:3, 1:3, ~ .x + .y)
```

```{r}
# If you need more than two arguments, you can use .n for the nth argument:
list(1:3, 1:3, 1:3) %>% pmap_dbl(~ .1 + .2 + .3) # or ..1 ..2 ..3
```

