---
title: "R4DS-QuickRef -- Ch3 Representing Tables: tibble"
output: html_notebook
---

```{r}
library(tibble)
```

### Creating tibbles

Tidyverse functions that create tabular data create tubbles rather than data frames.
```{r}
x <- readr::read_csv("../import/data/data.csv")
```

```{r}
x
```
tibbles have several superclasses including "data.frame", meaning that you can most
often use tibbles in place of data frames. But not always!
```{r}
class(x)
```

```{r}
# Create a data frame from tibble
(y <- as.data.frame(x))
```

```{r}
# Create a tibble from a data frame
(z <- tibble::as_tibble(y))
```

You can create a tibble from vectors using `tibble()`. If the tibble has more than
20 lines you will see only the first 10. The second thing to notice is that you can 
refer to previous columns when specifying later columns.
```{r}
(xx <- 
   tibble(
  x = 1:100,
  y = x^2,
  z = y^2
   )
)
```

You can also define a tibble one line per row with `tribble()´
```{r}
tribble(
  ~x, ~y, ~z,
  1, 1, 1,
  2, 4, 16,
  3,  9, 81
)
```

Remember that ´readr::read_csv` can also be used to create tibbles from strings
```{r}
readr::read_csv(
  "x, y, z
  1, 1, 1
  2, 4, 16
  3, 9, 81
  "
)
```

### Indexing Tibbles

```{r}
(x <- readr::read_csv("../import/data/data.csv"))
```

```{r}
(y <- as.data.frame(x))
```
Single-brackets extract a tibble or a df, respectively
```{r}
x["A"]
```

```{r}
y["A"]
```
```{r}
x[1]
```

```{r}
y[1]
```
Double-brackets and $ extract a vector from both tibbles and data frames
```{r}
x[[1]]
x$A
y[[1]]
y$A
```

More than one column with []
```{r}
x[c("A", "B")]
x[c(1, 2)]
y[2:3]
y[c("B", "C")]
```

Extract a subset of the tibble using rows and columns
```{r}
x[1:2, 2:3]
y[1:2, 2:3]
```

Subset of rows
```{r}
x[1:2, ]
y[2:3, ]
```

Subset of columns
```{r}
x[, 2:3]
y[, c("A", "C")]
```

If you extract a subset of rows from a single column, tibbles and data frames no 
longer have the same behavior. A tibble will give you a tibble in return, while a data 
frame will give you a vector:
```{r}
# Produces a tibble
x[1:2, 2]
```

```{r}
# Produces a vector
y[1:2, 2]
```