---
title: "Plotting"
output: html_notebook
---

```{r}
suppressPackageStartupMessages({
  library(tidyverse)
  library(palmerpenguins) 
})
```

```{r}
penguins
```

### Capabilities in ggplot2

```{r}
penguins %>% 
  ggplot()
```

```{r}
penguins %>% 
  ggplot(aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point()
  
```

```{r}
penguins %>% 
  ggplot(aes(x = flipper_length_mm, y = body_mass_g, color = sex)) +
  geom_point()
```

```{r}
penguins %>% 
  ggplot(aes(x = flipper_length_mm, y = body_mass_g, color = sex)) +
  geom_point() +
  scale_color_manual(values = c("darkorange", "cyan4"), na.translate = FALSE) +
  cowplot::theme_cowplot()
```

```{r}
penguins %>% 
  ggplot(aes(x = flipper_length_mm, y = body_mass_g, color = sex)) +
  geom_point() +
  scale_color_manual(values = c("darkorange", "cyan4"), na.translate = FALSE) +
  cowplot::theme_cowplot() +
  labs(
    title = "Penguin flipper and body mass",
    caption = "Horst AM, Hill AP, Gorman KB (2020)",
    subtitle = 
      "Dimensions for Adelie, Chinstrap and Gentoo Penguins at Palmer Station LTER"
  )
```


```{r}
penguins %>% 
  ggplot(aes(x = flipper_length_mm, y = body_mass_g, color = sex)) +
  geom_point() +
  scale_color_manual(values = c("darkorange", "cyan4"), na.translate = FALSE) +
  cowplot::theme_cowplot() +
  labs(
    title = "Penguin flipper and body mass",
    caption = "Horst AM, Hill AP, Gorman KB (2020)",
    subtitle = 
      "Dimensions for Adelie, Chinstrap and Gentoo Penguins at Palmer Station LTER"
  ) +
  theme(plot.subtitle = element_text(size = 12))
```

 
```{r}
penguins %>% 
  ggplot(aes(x = flipper_length_mm, y = body_mass_g, color = sex)) +
  geom_point() +
  scale_color_manual(values = c("darkorange", "cyan4"), na.translate = FALSE) +
  cowplot::theme_cowplot() +
  labs(
    x = "Flipper length (mm)",
    y = "Body mass (g)",
    color = "Penguin sex",
    title = "Penguin flipper and body mass",
    caption = "Horst AM, Hill AP, Gorman KB (2020)",
    subtitle = "Dimensions for Adelie/Chinstrap/Gentoo Penguins at Palmer Station LTER"
  ) +
  theme(plot.subtitle = element_text(size = 12)) 
```

```{r}
penguins %>% 
  ggplot(aes(x = flipper_length_mm, y = body_mass_g, color = sex)) +
  geom_point() +
  scale_color_manual(values = c("darkorange", "cyan4"), na.translate = FALSE) +
  cowplot::theme_cowplot() +
  labs(
    x = "Flipper length (mm)",
    y = "Body mass (g)",
    color = "Penguin sex",
    title = "Penguin flipper and body mass",
    caption = "Horst AM, Hill AP, Gorman KB (2020)",
    subtitle = "Dimensions for Adelie/Chinstrap/Gentoo Penguins at Palmer Station LTER"
  ) +
  theme(
    plot.subtitle = element_text(size = 12),
    legend.position = "bottom",
    plot.caption = element_text(hjust = 0, face = "italic"),
    plot.caption.position = "plot"
  )

```

```{r}
penguins %>% 
  ggplot(aes(x = flipper_length_mm, y = body_mass_g, color = sex)) +
  geom_point() +
  scale_color_manual(values = c("darkorange", "cyan4"), na.translate = FALSE) +
  cowplot::theme_cowplot() +
  labs(
    x = "Flipper length (mm)",
    y = "Body mass (g)",
    color = "Penguin sex",
    title = "Penguin flipper and body mass",
    caption = "Horst AM, Hill AP, Gorman KB (2020)",
    subtitle = "Dimensions for Adelie/Chinstrap/Gentoo Penguins at Palmer Station LTER"
  ) +
  theme(
    plot.subtitle = element_text(size = 12),
    legend.position = "bottom",
    plot.caption = element_text(hjust = 0, face = "italic"),
    plot.caption.position = "plot"
  ) +
  facet_wrap(~ species)

```
 
 
```{r}
penguins %>% 
  ggplot(aes(x = flipper_length_mm, y = body_mass_g, color = sex)) +
  geom_point() +
  scale_color_manual(values = c("darkorange", "cyan4"), na.translate = FALSE) +
  cowplot::theme_cowplot() +
  labs(
    x = "Flipper length (mm)",
    y = "Body mass (g)",
    color = "Penguin sex",
    title = "Penguin flipper and body mass",
    caption = "Horst AM, Hill AP, Gorman KB (2020)",
    subtitle = "Dimensions for Adelie/Chinstrap/Gentoo Penguins at Palmer Station LTER"
  ) +
  theme(
    plot.subtitle = element_text(size = 12),
    legend.position = "bottom",
    plot.caption = element_text(hjust = 0, face = "italic"),
    plot.caption.position = "plot"
  ) +
  facet_wrap(~ species) +
  scale_x_continuous(guide = guide_axis(n.dodge = 2))
```
 
```{r}
penguins %>% 
  ggplot(aes(x = flipper_length_mm, y = body_mass_g, color = sex)) +
  geom_point() +
  scale_color_manual(values = c("darkorange", "cyan4"), na.translate = FALSE) +
  cowplot::theme_cowplot() +
  labs(
    x = "Flipper length (mm)",
    y = "Body mass (g)",
    color = "Penguin sex",
    title = "Penguin flipper and body mass",
    caption = "Horst AM, Hill AP, Gorman KB (2020)",
    subtitle = "Dimensions for Adelie/Chinstrap/Gentoo Penguins at Palmer Station LTER"
  ) +
  theme(
    plot.subtitle = element_text(size = 12),
    legend.position = "bottom",
    plot.caption = element_text(hjust = 0, face = "italic"),
    plot.caption.position = "plot"
  ) +
  facet_wrap(~ species) +
  scale_x_continuous(guide = guide_axis(n.dodge = 2)) +
  scale_y_continuous(labels = scales::label_comma())
```

 ### Your first plot
 
```{r}
# Note that aes is specified inside the geom_point here
penguins %>% 
  ggplot() + 
  geom_point(aes(bill_length_mm, bill_depth_mm, color = species))
```


```{r}
# Refer to one of the column data, here: mistake
penguins %>% 
  ggplot() + 
  geom_point(aes(bill_length_mm, bill_depth_mm, color = country))
```


```{r}
# Passing the unknown column as string as a different effect:
# This is hardly useful, but we shall see an application later.
# Stick to the 2 mapping rules: In aes() Refer to a valid column.

penguins %>% 
  ggplot() + 
  geom_point(aes(bill_length_mm, bill_depth_mm, color = "country"))
```

Mapping aesthetics correctly
In aes() and refer to a data column

```{r}
penguins %>% 
  ggplot() +
  geom_bar(aes(y = species,
               fill = sex))
```

How not using a string for mapping is useful?

Fair question

    Could we pass an expression?
    Which penguins are above 4 kg?
    Use body_mass_g > 4000 that return a boolean to find out

```{r}
# The expression was evaluated in penguins context 
# Obvious that Gentoo are bigger than the 2 other species
ggplot(penguins) +
  geom_bar(aes(y = species,
               fill = body_mass_g > 4000))
```

Inheritance of arguments across layers

Compare the two following (great example of a Simpson's paradox):


    aesthetics in ggplot() are passed on to all geometries.
    aesthetics in geom_*() are specific (and can overwrite inherited)

```{r}
ggplot(penguins,
       aes(x = bill_length_mm,
           y = bill_depth_mm)) +
  geom_point(aes(colour = species)) +
  geom_smooth(method = "lm", formula = 'y ~ x')
```

```{r}
ggplot(penguins,
       aes(x = bill_length_mm,
           y = bill_depth_mm,
           colour = species)) +
  geom_point() +
  geom_smooth(method = "lm", formula = 'y ~ x')
```

Try it

    Map the island variable to a shape aesthetics for both dots and linear models
    All dots (circles / triangles / squares) with:
        A size of 5
        A transparency of 30% (alpha = 0.7)

```{r}
penguins %>% 
  ggplot(aes(bill_length_mm, bill_depth_mm, shape = island, color = species)) +
  geom_point(size = 5, alpha = 0.7) +
  geom_smooth(method = "lm", formula = 'y ~ x')
```

### Joining observations

```{r}
set.seed(212) # tidyr::crossing generate combinations
tib <- tibble(crossing(x = letters[1:4], 
                       g = factor(1:2)), 
              y = rnorm(8))
tib
```

Invisible aesthetic: grouping