---
title: "R4DS-QR QuickRef: Ch 10 -- Working with Factors: forcats"
output: html_notebook
---

```{r}
library(tidyverse)
```

#### Creating factors
```{r}
# If you don't specify levels, you get one per unique element
factor(c("A", "B", "C"))
```

```{r}
# Levels defined explicitly
factor(c("A", "B", "C"), levels = c("A", "B", "C", "D"))
```


```{r}
# Levels defined explicitly in different order
factor(c("A", "C", "B"), levels = c("D", "B", "C", "A"))
```

```{r}
# I you have an element not defined in the levels, you get NA
factor(c("A", "B", "C", "D"), levels = c("A", "B", "C"))
```

The factor() function is usually the right choice for creating factors. The function 
is not generic, however, so if you want a way to translate a data structure you have 
implemented yourself, one that needs a specific method to do so, then you might want to 
specialize as_factor() from forcats.

If you specialize as_factor(), you can do anything you want, but on normal vectors, 
as_factor() works similarly to as.factor() from base R. The two functions differ, 
though, in how they set levels. If you use as.factor(), the levels will be sorted.

```{r}
# Levels ordered alphabetically
(f1 <- as.factor(c("C", "B", "B", "A", "D")))
```

```{r}
# Order is as it is given
(f2 <- as_factor(c("C", "B", "B", "A", "D")))
```

#### Concatenation

If you have two factors, then you can concatenate them using fct_c(). The levels of the 
resulting factor are the union of the levels of the input factors. The order of the levels 
depends on the order of the input factors. The levels in the first factor go first, then 
the levels in the second. If the factors share levels, then the order is determined by the 
first vector

```{r}
f1
f2
```

```{r}
# The order of the factors depends on the order they are concatenated
fct_c(f1, f2)
fct_c(f2, f1)
```

```{r}
# A factor with levels not found in f1
(f3 <- as_factor(c("X", "Y", "A")))
```

```{r}
# If f1 is first, the unique levels of f3 are included at the end
fct_c(f1, f3)
```

```{r}
# If f3 is first, f3 levels come first
fct_c(f3, f1)
```

```{r}
# You can concatenate more than two factor vectors. The previous rules apply
fct_c(f3, f2, f1)
```

If you have several factors that do not have the same levels, but you want them to 
have (without concatenating them), you can use fct_unify(). It takes a list of factors 
and gives you a list of factors where they all have as levels the union of the levels of 
the input.

```{r}
fct_unify(list(f1, f2, f3))
```

#### Projection

It happens that your categorical data is too fine-grained and you want to group 
categories into larger classes. If so, the function you want is fct_collapse().
It lets you  map your existing levels to new levels. Its first argument is a 
factor, and after that, you  provide named arguments. As parameters to the 
named arguments, you must provide a list of level names. Each name becomes 
a level, and that level will contain the elements in the list you give as the argument.

```{r}
fct_collapse(
  fct_c(f3, f1),
  a = c("A", "X"),
  b = c("B", "Y"),
  c = c("C", "D")
)
```

```{r}
# You do not need to remap all levels. Those you do not map will stay as they were.
fct_collapse(
  fct_c(f3, f1),
  a = c("A", "X"),
  b = c("B", "Y"),
)
```

If you only want to rename and not collapse levels, you can use fct_recode().

```{r}
# Renaming some levels
f1
fct_recode(f1, a = "A", b = "B")
```

```{r}
# Using fct_recode or fct_collapse
fct_recode(f1, X = "A", X = "B", Y = "C", Y = "D")
fct_collapse(f1, X = c("A", "B"), Y = c("C", "D"))
```

If you want to reduce the number of levels based on how many elements they have, 
rather than merge levels in an explicitly specified way, then you can use fct_lump(). You 
can merge levels such that you keep the n levels with the most elements, for example, for 
n = 5 and n = 2:

```{r}
f <- factor(sample(LETTERS[1:10], 20, replace = TRUE), levels = LETTERS[1:10])
f
table(f)
```

```{r}
f %>% fct_lump(n = 5, other_level = "X") %>% table()
```

```{r}
f %>% fct_lump(n = 2, other_level = "X") %>% table()
```

You see more than five and two levels here, even though we called fct_lump() with 
n = 5 and n = 2. If there are several levels with the same number of elements, the 
function doesn’t pick random ones to keep. You get all with the highest occurrences.

Instead of picking the number of categories you want, you can require that they 
contain at least or at most a fraction of the data. 

```{r}
f %>% fct_lump(prop = 0.1, other_level = "X") %>% table()
```

If you want the categories that contain less than 10% of the data, you use prop as well 
but with a negative value.

```{r}
f %>% fct_lump(prop = -0.1, other_level = "X") %>% table()
```

The fct_other() lets you pick a list of categories to keep or drop. If you 
keep categories, then all the others are put in the same “other” category. If you drop 
categories, then those you drop are placed in an “other” category. You can name the 
“other” category using the argument other_level.

```{r}
f %>% fct_other(keep = LETTERS[1:5], other_level = "X")
```

Sometimes, what you want is even simpler. You have levels that are not found in the 
actual data. You can get rid of the empty levels using fct_drop().

```{r}
f1
levels(f1) <- LETTERS[1:10]
```

```{r}
fct_drop(f1)
```

#### Adding Levels

With the fct_expand() function, you can add levels that are not in your factor, and 
any levels that are already in the data will be ignored.

```{r}
f1
f3
```

```{r}
fct_expand(f1, levels(f3))
```

You can use fct_na_value_to_level() to map all missing data into a new 
category:

```{r}
f2
```

```{r}
fna <- f2
fna[2] <- NA
fna[3] <- NA
fna
```

```{r}
fct_na_value_to_level(fna, level = "Missing")
```

#### Reorder Levels

If you have a factor created with factor() without specifying the levels, then the 
levels will be all elements seen in the input sorted in their natural order.

```{r}
f <- factor(sample(LETTERS[1:5], 10, replace = TRUE))
f
```

The function fct_inorder() will order your levels to match the order that the 
categories are seen in the data—the order you get from as_factor().
```{r}
# Order is the order of appearance
fct_inorder(f)
```

The functions for reordering levels can also make the factors ordered.

```{r}
# Note the odered levels
fct_inorder(f, ordered = TRUE)
```

If you want to sort the levels, you do not need forcats functions. The plain old 
factor() will do nicely.

```{r}
# Base R does the trick
factor(f, levels = sort(levels(f)))
```

```{r}
factor(f, ordered = TRUE)
factor(f, levels = rev(levels(f)), ordered = TRUE)
```

You can order levels by their frequency in the factor using fct_infreq():

```{r}
# Using fct_count and some dplyr to get a table of freqs
fct_count(f) %>% arrange(desc(n))
```

```{r}
# See how fct_infreq sorts the levels
# Order and reverse oder
fct_infreq(f)
fct_infreq(f, ordered = TRUE)
fct_rev(fct_infreq(f, ordered = TRUE))
```

In the last example, I reversed the levels, so the order is smallest to largest rather 
than largest to smallest. The function fct_rev() reverses the levels; a frequent use for 
this is when you plot with an axis given by the factor. There, the order is given by the 
levels, and you might want it reversed.





