---
title: "R4DS-Quickref: Ch6 -- Pipelines: magrittr"
output: html_notebook
---

```{r}
library(tidyverse)
```

### Pipeline Notation

x %>% f() %>% g() %>% h()

or

x |> f() |> g() |> h()

is equivalent to

h(g(f(x)))

So, by the way, is

x %>% f %>% g %>% h

but not

x |> f |> g |> h

The native operator needs to have a function call, f(), on the right-hand side of the 
operator, while the magrittr will accept either a function call, f(), or just a function 
name, x %>% f. You can also give %>% an anonymous function:

```{r}
# You must put the lambda-fn in parentheses with both %>%  and |>
"hello, "%>% (\(y) paste(y, "world!"))
```

but |> still needs a function call, so if you give it an anonymous function, you must 
call the function as well as add () after the function definition).

```{r}
# with |> you have to call the function at the end with ()
"hello, "|> (\(y) paste(y, "world!")) () 
```

With the %>% operator, you do not need the parentheses for the function calls, but 
most prefer them to make it clear that we are dealing with functions, and the syntax is 
then still valid if you change the code to the |> operator. Also, if your functions take more 
than one argument, parentheses are needed, so if you always include them, it gives you a 
consistent notation.

### Pipelines and Function Arguments

With the magrittr operator, %>%, but not the |> operator, you can change the default 
input position using the special variable . (dot). If the right-hand side expression has . 
as a parameter, then that is where the left-hand side value goes.

```{r}
# You can use a dot more than once on the right-hand side:
rnorm(5) %>% tibble(x = ., y = .)
```


```{r}
# You can also use it in nested function calls.
rnorm(5) %>% tibble(x = ., y = abs(.))
```

```{r}
# If the dot is only found in nested function calls, however, magrittr will still add it as 
# the first argument to the right-hand side function.
rnorm(5) %>% tibble(x = sin(.), y = abs(.))
```

```{r}
# You can avoid this by putting the rhs expression in curly brackets
rnorm(5) %>% {tibble(x = sin(.), y = abs(.))}
```

In general, you can put expressions in curly brackets as the right-hand side of a pipe 
operator and have magrittr evaluate them. Think of it as a way to write one-parameter 
anonymous functions. The input variable is the dot, and the expression inside the curly 
brackets is the body of the function.

### Function Composition

Many functional programming languages encourage you to write functions by 
combining other functions and have operators for that. It is not frequently done in R, 
and while you can implement function composition, there is no built-in operator. The 
magrittr package, however, gives you this syntax:

h <- . %>% f() %>% g()

This defines the function h, such that h(x) = g(f(x)).
If we take the tidy-then-clean pipeline we saw earlier, and imagine that we need to 
do the same for several input files, we can define a pipeline function for this as
```{r}
# Compose function pipeline
pipeline <- 
  . %>%
 pivot_longer(
   names_to = "year", 
   values_to = "mean_income", 
   !country) %>%
 drop_na()

pipeline
```


Note that purrr has a `compose()` function to, well, compose functions. This can 
be combined with `partial()`.

```{r}
# Basic composition goes left<-right
fn <- 
  compose(
    \(x) paste(x, "foo"), 
    \(x) paste(x, "bar")
  )

fn("input")
```

```{r}
# Change direction with .dir 
fn2 <- 
  compose(
    \(x) paste(x, "foo"), 
    \(x) paste(x, "bar"), 
    .dir = "forward"
  )

fn2("input")
```

```{r}
# Create partial functions with set parameters
# Changing the order of arguments if quite funky

add_foo1 <- partial(paste, "foo") # Default, partialised arguments first
add_foo2 <- partial(paste, ... =, "foo") # Partialised arguments last

add_foo1("bar")  
add_foo2("bar")
```


```{r}
# Using compose() with partial()
add_foo <- partial(paste, ... =, "foo")
add_bar <- partial(paste, ... =, "bar")

fn3 <- compose(add_foo, add_bar)
fn4 <- compose(add_foo, add_bar, .dir = "forward")

fn3("input")
fn4("input")
```

### Other Pipe Operations
There are three other pipe operators in magrittr. These are not imported when you 
import the tidyverse package, so to get access to them, you have to import magrittr
explicitly.

```{r}
library(magrittr)
```
`%<>%` is the assignment pipe. This reassignment operator behaves similarly to the stepwise pipeline convention we considered at the start of the chapter, but it makes explicit that we are updating an 
existing variable. You cannot accidentally assign the result to a different variable. 
```{r}
# These two are equivalent
(my_letters <- letters[1:5])
my_letters <- my_letters %>% toupper()
my_letters

(my_other_letters <- letters[6:10])
my_other_letters %<>% toupper()
my_other_letters
```

If your right-hand side is an expression in curly brackets, you can refer to the input 
through the dot variable:
```{r}
(mydata <- tibble(x = rnorm(5), y = rnorm(5)))
```
```{r}
mydata %>% { .$x - .$y }
```

If, as here, the input is a data frame and you want to access its columns, you need the 
notation .$ to access them. The `%$%` pipe operator opens the data frame for you so you 
can refer to the columns by name.

```{r}
mydata %$% { x - y } 
```

The tee pipe operator, `%T>%`, behaves like the regular pipe operator, %>%, except that it 
returns its input rather than the result of calling the function on its right-hand side. The 
regular pipe operation x %>% f() will return f(x), but the tee pipe operation x %T>% 
f() will call f(x) but return x. This is useful for calling functions with side effects as a 
step inside a pipeline, such as saving intermediate results to files or plotting data.

If you call a function in a usual pipeline, you will pass the result of the function call 
on to the next function in the pipeline. If you, for example, want to plot intermediate 
data, you might not be so lucky that the plotting function will return the data. The 
ggplot2 functions will not. If you want to plot intermediate values, you need to save 
the data in a variable, save it, and then start a new pipeline with the data as input.

```{r}
# Using %T>% to print at the middle of the pipeline 
iris %T>% 
  { select(., Sepal.Length) %>% print(head) } %>%  # Print to screen and then continue with pipeline
  pull(Species) %>% 
  toupper()
```