---
title: "Data Wrangling 1"
output: html_notebook
---

```{r}
library(tidyverse)
```

```{r}
(judgments <- read_tsv("data/judgments.tsv"))
```

```{r}
# Ugly and convoluted, first column is a factor! 
as.data.frame(table(judgments[["condition"]]))
```

```{r}
# Better way
judgments %>% 
  count(condition)
```

```{r}
# Sort results
judgments %>% 
  count(condition, sort = TRUE)
```

```{r}
# Previous is a shortcut for grouping and summarising
judgments %>% 
  group_by(condition) %>% 
  summarise(n = n(), .groups = "drop") # The dropping is only required when several groupings
```

```{r}
# Previous iusing the .by version of grouping
judgments %>% 
  summarise(n = n(), .by = condition)
```

When counting is not enough


    summarise returns as many rows as groups - one if no groups.
    mutate returns as many rows as given.

```{r}
judgments %>% 
  summarise(
    min = min(mood_pre, na.rm = TRUE), 
    max = max(mood_pre, na.rm = TRUE)
    )
```

```{r}
judgments %>%
  mutate(
    min = min(mood_pre, na.rm = TRUE), 
    max = max(mood_pre, na.rm = TRUE), .before = 1
    )
```

If we want the min/max per condition
```{r}
judgments %>% 
  summarise(
    min = min(mood_pre, na.rm = TRUE), 
    max = max(mood_pre, na.rm = TRUE),
    .by = condition # or group_by(condition)
    )
```

```{r}
judgments %>% 
  mutate(
    min = min(mood_pre, na.rm = TRUE), 
    max = max(mood_pre, na.rm = TRUE), 
    .by = condition, 
    .before = 1
  )
```

Within one summarise statement
```{r}
judgments %>% 
  summarise(
    n_rows = n(), 
    n_subject = n_distinct(subject),
    first_id = first(subject),
    last_id = last(subject),
    mean = mean(mood_pre, na.rm= TRUE),
    id_10 = nth(subject, n = 10)
  )
```

group_by() results in a persistent group

    Each call to the tibble will respect the grouping
    Summarize will strip away one level of grouping

Grouping by more than one variable -> Peeling effect

For variable grouping, one is peeled off from the right 

Ask explicitly to ungroup data

    by ungroup()
    by .groups argument to keep or drop groups.

```{r}
# One grouping var remains
judgments %>% 
  group_by(condition, gender) %>% 
  summarise(n_ans = n_distinct(STAI_post_1_1)) %>% 
  group_vars(
  )
```

```{r}
# All groupings dropped
judgments %>% 
  group_by(condition, gender) %>%  
  summarise(n_ans = n_distinct(STAI_post_1_1),
            .groups = "drop") %>% 
  group_vars()
```


```{r}
# Or, use ungroup to drop remaining group vars
judgments %>% 
  group_by(condition, gender) %>%  
  summarise(n_ans = n_distinct(STAI_post_1_1)) %>% 
  ungroup() %>% 
  group_vars()
```


Arranging values within groups
arrange() can sort values by multiple columns

```{r}
# Arrange ignores groupings!
judgments %>%  
  mutate(mood_pre_cat = case_when(
    mood_pre < 25  ~ "poor", 
    mood_pre > 75 ~ "great",
    .default = "normal")) %>% 
  group_by(mood_pre_cat) %>%  
  arrange(desc(mood_post)) %>%  
  select(mood_pre_cat, mood_post) %>%  
  distinct()
```

For arrange() and grouping

Solution -> Use .by_group = TRUE

```{r}
# use .by_group to arrange by groupings
judgments %>%  
  mutate(mood_pre_cat = case_when(
    mood_pre < 25  ~ "poor", 
    mood_pre > 75 ~ "great",
    .default = "normal")) %>% 
  group_by(mood_pre_cat) %>%  
  arrange(desc(mood_post), .by_group = TRUE) %>%  
  select(mood_pre_cat, mood_post) %>%  
  distinct()
```

But, you are better off using arrange with two variables, like below
```{r}
judgments %>%  
  mutate(mood_pre_cat = case_when(
    mood_pre < 25  ~ "poor", 
    mood_pre > 75 ~ "great",
    .default = "normal")) %>% 
  group_by(mood_pre_cat) %>%  
  arrange(mood_pre_cat, desc(mood_post)) %>%  
  select(mood_pre_cat, mood_post) %>%  
  distinct()
```

Dealing with multiple return values per group

    range() returns min and max
    summarise() duplicates the key names

```{r}
# Summarise keeps groupings here
judgments %>% 
  group_by(condition, gender) %>% 
  summarise(range = range(mood_pre, na.rm = TRUE), n = n()) 
```

"Warning: Returning more (or less) than 1 row per `summarise()` group was deprecated in dplyr 1.1.0.
Please use `reframe()` instead.
When switching from `summarise()` to `reframe()`, remember that `reframe()` always returns an ungrouped data frame and adjust accordingly."

Using `reframe`

reframe() creates a new data frame by applying functions to columns of an existing data frame. It is most similar to summarise(), with two big differences:

    reframe() can return an arbitrary number of rows per group, while summarise() reduces each group        down to a single row.

    reframe() always returns an ungrouped data frame, while summarise() might return a grouped or           rowwise data frame, depending on the scenario.

We expect that you'll use summarise() much more often than reframe(), but reframe() can be particularly helpful when you need to apply a complex function that doesn't return a single summary value. 

```{r}
judgments %>% 
  group_by(condition, gender) %>% 
  reframe(range = range(mood_pre, na.rm = TRUE), n = n())
```

More advanced but useful: 3 quantiles

```{r}
# Summarise three values per group
judgments %>% 
  filter(!is.na(mood_pre)) %>% 
  group_by(condition, gender) %>% 
  reframe(
    quan = quantile(mood_pre, c(0.25, 0.5, 0.75)),
    q = c(0.25, 0.5, 0.75),
    n = n())
```

Grouping and mutate()

```{r}
judgments %>% 
  group_by(condition) %>% 
  mutate(n = n()) %>% 
  relocate(condition, n)
```

Use judgments to compute basic statistics for all moral dilemma columns considering the conditions:

    Compute the mean, the median, the standard deviation as well as min and max values.
    Find meaningful short names for the functions such as med for median().
    Assign judgments_condition_stats to the results.

```{r}
judgments %>% 
  summarise(
    across(contains("dilemma"), 
           list(mean = mean,
                median = median,
                sd = sd,
                min = min,
                max = max),
           .names = "{.fn}.{.col}")
  )
```

In judgments:

    Find the number of subjects by age, gender and condition, e.g. how many 20 years of age females are     in the stress group.
    
    Sort the resulting tibble such that the condition that contains the most populous group is sorted       first (i.e. stress or control appear together).
    Ensure that the resulting tibble does not contain groups.

```{r}
# Using .by
judgments %>%
  summarise(n = n(), .by = c(condition, gender, age)) %>% 
  arrange(condition, gender, desc(n)) # manully pealed off age
```


```{r}
# Using normal group_by
judgments %>% 
  group_by(condition, gender, age) %>% 
  summarise(n = n()) %>% 
  arrange(desc(n), .by_group = TRUE) %>% 
  ungroup()
```


































