---
title: "Joins"
output: html_notebook
---

```{r}
# Load tidyverse
suppressPackageStartupMessages(
  library(tidyverse)
)
```

### Matching subjects in two tibbles

```{r}
judgments <- read_tsv("data/judgments.tsv", show_col_types = FALSE)
head(judgments)
```



```{r}
# Additional data for some participants for coffee consumption
(coffee_drinkers <-
  tribble(
    ~student, ~coffee_shots,
    21,       1,
    23,       4,
    28,       2
  ))
```

```{r}
# Smaller sample set. This is made up data for demonstration purposes only.
(subject_mood <-
  judgments %>% 
  select(subject, condition, gender, starts_with("mood")) %>% 
  distinct())
  
```

#### Combining the two tables

```{r}
# Error! No common key.
inner_join(coffee_drinkers, subject_mood)
```

```{r}
# Provide the corresponding columns names.
inner_join(subject_mood, coffee_drinkers, by = join_by(subject == student))
```

Creating new tables through joins

    Key operations in data processing

    Role of observations as row changes

    inner_join() is the most strict join operations

    merge is a similar operation in base R


```{r}
# full_join example
subject_mood %>% 
  full_join(coffee_drinkers, by = join_by(subject == student))
```

Two tables - same column names

What if we have gender in both tables?

```{r}
# Add gender to coffee drinkers
(coffee_drinkers <- 
   mutate(coffee_drinkers, gender = c("female", "female", "male")))
```
Mutating joins distinct the same column names by adding suffixes - .x and .y.

```{r}
left_join(subject_mood,
          coffee_drinkers,
          by = join_by(subject == student)) %>% 
  arrange(gender.y)
```

Join by one column

```{r}
# Add suffixes
left_join(subject_mood,
          coffee_drinkers,
          by = join_by(subject == student), 
          suffix = c( "_mood", "_coffee")) %>% 
  arrange(gender_coffee)
```

Join by two columns
```{r}
# Now we get just one column for gender
left_join(subject_mood,
          coffee_drinkers,
          by = join_by(subject == student, gender)) %>% 
  arrange(coffee_shots)
```

Filtering joins -

    Only the existence of a match is important; it doesn’t matter which observation is matched. This means that filtering joins never duplicate rows like mutating joins do

semi_join()

Filter matches in x, no duplicates.

anti_join()

Extract what does not match.


semi_join() does not alter original
```{r}
(tx <- tribble(~key, ~x,
              1, "x1", 
              2, "x2", 
              3, "x3", 
              4, "x4"))
```

```{r}
(ty <- tribble(~key, ~y,
              1, "y1",
              2, "y2",
              2, "y3",
              3, "y4"))
```

```{r}
# Filtering with semi_join
semi_join(tx, ty, by = join_by(key))
```
```{r}
# Mutating with inner_join
inner_join(tx, ty, by = join_by(key))
```

Converting into long or wide formats - pivot functions

```{r}
(variants_wide <- 
  tribble(
    ~sample_id, ~`3`, ~`5`, ~`8`,
    "L1002",    "A",  "C",  NA,
    "L1034",    "A",  NA,   "T",
    "L1234",    NA,   "C",  "T"
  ))
```

```{r}
variants_wide %>% 
  pivot_longer(-sample_id, names_to = "pos", values_to = "variant")
```

Values across dilemmas

```{r}
# Select, pivot wider and count
  judgments %>% 
  select(subject, condition, age, starts_with("moral_dilemma")) %>% 
  pivot_longer(
    contains("dilemma"),
    names_prefix = "moral_dilemma_",   
    # names_pattern = "moral_dilemma_(.*)", # This is an alternative using regex
    names_to = "dilemma",
    values_to = "dilemma_val") %>% 
  count(condition, dilemma_val) 
```

```{r}
# Pivot wider
(judgments_condition_stats <-
  judgments %>% 
  select(subject, condition, age, starts_with("moral_dilemma")) %>% 
  pivot_longer(
    contains("dilemma"),
    names_prefix = "moral_dilemma_",   
    names_to = "dilemma",
    values_to = "dilemma_val") %>%
  count(condition, dilemma_val) %>% 
  # names_prefix is useful if names_from is a numeric vector 
  pivot_wider(names_from = dilemma_val, values_from = n, names_prefix = "score"))
```

Removing the data frame context through pull()

```{r}
# Extract one column from a df or tibble as a vector (useful in pipelines)
pull(judgments, age)[1:10]
```

```{r}
# Extract same vector along with a chosen name column
pull(judgments, age, name = gender)[1:10]
```

Calculate differences between subjects

```{r}
judgments %>% 
  select(subject, mood_pre) %>%  
  arrange(subject) %>%  
  mutate(prev_mood_pre = lag(mood_pre),
         mood_diff = mood_pre - lag(mood_pre)) 
```


```{r}
(input <- 
  tribble(
    ~sampleid, ~var1, ~var2, ~var3,
    "S1", 	"A3T", 	"T5G", 	"T6G",
    "S2", 	"A3G", 	"T5G", 	NA,
    "S3", 	"A3T", 	"T6C", 	"G10C",
    "S4", 	"A3T", 	"T6C", 	"G10C"
  ))
```
