---
title: "R4DS-Quickref: Ch5 -- Reformatting tables: tidyr"
output: html_notebook
---

```{r}
library(tidyverse)
```

### Tidy Data

Each column is a single variable, each row is a single observation 

Consider the data below:
```{r}
# Untidy data
(mean_income <-
   tribble(
     ~country,~ '2001',~ '2002',~ '2003',~ '2004',~ '2005',
     "Numenor", 123456, 132654, 321646, 324156, 325416,
     "Westeros", 314256, 432165, 546123, 465321, 561423,
     "Narnia", 432156, 342165, 564123, 543216, 465321,
     "Gondor", 531426, 321465, 235461, 463521, 561423,
     "Laputa", 14235, 34125, 45123, 51234, 54321
   )
)
```

### Pivoting
```{r}
(tidy_income <-
   mean_income %>% 
   pivot_longer(
     cols = c(`2001`, `2001`, `2002`,
              `2003`, `2004`, `2005`),
     names_to = "year",
     values_to = "income"
   )
)
```

```{r}
# Using : for choosing the columns works for contiguous cols
mean_income %>% 
  pivot_longer(
    cols = `2001`:`2005`,
    names_to = "year",
    values_to = "income"
  )
```

```{r}
# Using negation
mean_income %>% 
  pivot_longer(
    cols = -country,  # !country works too
    names_to = "year",
    values_to = "income"
  )
```

Generally, you can combine pivot_longer() with all the tidy selection functions to 
select columns, for example, you can use starts_with() or combine starts_with()
and complement, !, or anything else the tidy selection language allows.

```{r}
# Another way of dropping 'country' from pivoting
mean_income %>% 
  pivot_longer(
    cols = !starts_with("c"), # or, -starts_with("c") 
    names_to = "year",
    values_to = "income"
  )
```

Note that the `year` column has type `<chr>` above. If we want to assign a different 
type, we can use `names.transform`

```{r}
# Make 'year' an integer
mean_income %>% 
  pivot_longer(
    cols = -country,
    names_to = "year",
    values_to = "income",
    names_transform = as.integer
  )
```

```{r}
# 'year' as factor
mean_income %>% 
  pivot_longer(
    cols = -country,
    names_to = "year",
    values_to = "income",
    names_transform = as.factor
  )
```

There is a similar `values_transform` argument if you need to transform the values 
that go into the `values_to` column.

```{r}
# Make 'year' a factor and 'income' a character
mean_income %>% 
  pivot_longer(
    cols = -country,
    names_to = "year",
    values_to = "income",
    names_transform = as.factor,
    values_transform = as.character
  )
```

Sometimes you do need a non-tidy (wide) format. You can pivot a tidy dataset
wider by using `pivot_wider`

```{r}
tidy_income
```

```{r}
# Pivot wider
tidy_income %>% 
  pivot_wider(
    names_from = year,
    values_from = income
  )
```

### Complex Column Encodings

The best representation of a date is, of course, a date object, 
but for the sake of the example, let us say that we want to split a date into a day and a 
month column. You can do this using the `separate()` function.

```{r}
# A table
(tbl<-
   tribble(
     ~date,
     "11/5",
     "4/7",
     "21/12"
   )
)
```

```{r}
# Separate
# Default is to drop the original column
tbl %>% 
  separate(
    date, 
    into = c("day", "month"),
    sep = "/" 
  )
```

```{r}
# New syntax
tbl %>% 
  separate_wider_delim(
    date, 
    delim = "/", 
    names = c("day", "month")
  )
```

```{r}
# Old syntax: default separator is any non-alphanumeric character
tbl %>% 
  separate(
    date, 
    into = c("day", "month")
  )
```

```{r}
# Keep original column
# Note that all columns are <chr>
tbl %>% 
  separate(
    date,
    into = c("day", "month"),
    sep = "/",
    remove = FALSE
  )
```

```{r}
tbl %>% 
  separate_wider_delim(
    date, delim = "/", 
    names = c("day", "month"), 
    cols_remove = FALSE
  )
```

```{r}
# Convert new columns to more appropriate type
tbl %>% 
  separate(
    date,
    into = c("day", "month"),
    sep = "/",
    remove = FALSE,
    convert = TRUE
  )
```

```{r}
# No option to 'convert' in new syntax
tbl %>% 
  separate_wider_delim(
    date, delim = "/", 
    names = c("day", "month"), 
    cols_remove = FALSE
  )
```


If your separator is more complex, you can use the sep argument. The argument takes a regular 
expression, and the column will be split on this. For example, we can split on alphabet 
characters and spaces like this:

```{r}
# Complex separator
(tbl <- 
   tribble(
     ~date,
     "11th of month 5",
     "4th of month 7",
     "21st of month 12"
   )
)
```
```{r}
tbl %>% 
  separate(date, into = c("day", "month"), sep = "[[:alpha:][:space:]]+")
```
```{r}
# OR with separate_wider_regex
tbl %>% 
  separate_wider_regex(
    date, 
    patterns = c(day = "\\d+", "\\D+", month = "\\d+")
  )
```

Here is another unorthodox format
```{r}
(tbl <-
   tribble(
     ~date,
     "11th of May",
     "4th of July",
     "21st of December"
   )
)
```
```{r}
# separate_rows_wider superseeds 'extract'
tbl %>% 
  separate_wider_regex(
    date, 
    patterns = c(day = "\\d+", fluff = "[^[:upper:]]+", month = "\\w+")
  )
```

The reverse of separating is uniting.

```{r}
# Our trusted data
(tbl <- 
   tribble(
     ~date,
     "11/5",
     "4/7",
     "21/12"
   )
)
```

```{r}
# First we separate
(tbl2 <-
   tbl %>%
   separate_wider_delim(date, delim = "/", names = c("day", "month"))
)
```


```{r}
# Then we unite
# Default separator will be "_"
tbl2 %>% unite(date2, day:month)
```

```{r}
# Better like this
tbl2 %>% unite(date, day:month, sep = "/")
```
```{r}
# Set remove = FALSE, if you want to keep the original columns
tbl2 %>% unite(date, day:month, sep = "/", remove = FALSE)
```
Columns that contain more than one value do not always contain the same number 
of values. For example, we could have data such as the number of casualties per major 
group in WW1 and WW2.

```{r}
# WWI WWII casualities
(military_casualties <-
  tribble(
    ~war,~groups,~deaths,
    'WW1',
    "Allied Powers/Central Powers",
    "5.7,4.0",
    'WW2',
    "Germany/Japan/USSR/British Empire/USA",
    "5.3,2.1,10.7,0.6,0.4"
  )
 )
```

The groupings were not the same in the two wars, so we cannot split the data into 
different columns. We can, however, divide it into more rows using `separate_rows()`.
```{r}
# separate rows
military_casualties %>% 
  separate_rows(
    c(groups, deaths),
    sep = "/|," # / OR ,
  )
```

```{r}
# With separate_longer_delim use stringr::regex if you have multiple delims
military_casualties %>% 
  separate_longer_delim(cols = c(groups, deaths), delim = regex("[/,]")) 
```

```{r}
# Convers deaths to double
military_casualties %>% 
  separate_rows(
    c(groups, deaths),
    sep = "/|,",          # / OR ,
    convert = TRUE
  )
```

### Expanding, Crossing, and Completing

You can use `expand()` to get all the combinations of values, even the ones
that are missing from the data.

```{r}
# Generate data
(tbl <- 
   tribble(
     ~A,~B,~C,
     1,11,21,
     2,11,22,
     4,13,32
   )
)
```

```{r}
# Expand to all possible combinations of A, and B
# This is like expand_grid but starting with df columns. Only given columns used
tbl %>% expand(A, B)
```

```{r}
# You can also give a vector of values
tbl %>% expand(A = 1:4, B)
```

```{r}
# If you have vectors of values you want to combine this way, you can also use the crossing().
crossing(A = 1:4, B = 11:13)
```

```{r}
# If you only want the combinations found in the data, combine expand() with nesting()
(tbl <- 
   tribble(
     ~A, ~B, ~C,
     1, 11, 21,
     2, 11, 22,
     2, 11, 12,
     4, 13, 42,
     4, 13, 32
   )
 )
```

```{r}
tbl %>% expand(nesting(A, B))
```

```{r}
# If you need all combinations in a data frame, but accept missing values for those 
# that are not present, you can use the complete() function.

tbl %>% complete(A = 1:4)
```

```{r}
tbl %>% complete(B = 11:13)
```

```{r}
tbl %>% complete(A = 1:4, B = 11:13)
```

### Missing Values
```{r}
(mean_income <- 
   tribble(
     ~country, ~`2002`, ~`2003`, ~`2004`, ~`2005`,
     "Numenor", 123456, 132654, NA, 324156,
     "Westeros", 314256, NA, NA, 465321,
     "Narnia", 432156, NA, NA, NA,
     "Gondor", 531426, 321465, 235461, 463521,
     "Laputa", 14235, 34125, 45123, 51234,
   )
)
```

The simplest way to handle missing values is to get rid of it. A crude approach is 
to remove all observations with one or more missing variable values. The drop_na()
function does exactly that.

```{r}
# We lose a country entirely if we have missing data for a single year!
mean_income %>% 
  drop_na()
```

If we reformat the table and drop missing values, we will only remove the observations where 
the income is missing for a given year.

```{r}
# Pivot longer first
mean_income %>%
  pivot_longer(
    -country, 
    names_to = "year", 
    values_to = "income"
    ) 
```

```{r}
# Then drop NA's
mean_income %>%
  pivot_longer(
    -country, 
    names_to = "year", 
    values_to = "income"
    ) %>% 
  drop_na(income)
```

```{r}
# Remove NA's with pivot_longer
mean_income %>%
  pivot_longer(
    -country, 
    names_to = "year", 
    values_to = "income", 
    values_drop_na = TRUE
    )
```


Sometimes, you can replace missing data with some appropriate data. For example, 
we could replace missing data in the mean_income table by setting each value to the 
mean of its column. The replace_na() function does this. 

```{r}
mean_income %>% replace_na(list(`2003` = mean(.$`2003`, na.rm = TRUE)))
```

```{r}
mean_income %>%
  pivot_longer(
    -country, 
    names_to = "year", 
    values_to = "income", 
    ) %>% 
  replace_na(list(income = 0)) # Note, 0 is not a good replacement here, just an example!
```

A final function for managing missing data is the fill() function. It replaces missing 
values with the value above them in their column (or below them if you set the argument 
.direction to "up".

```{r}
# Generate some data
(tbl <- 
   read_csv(
     "year, quarter, income
 2011, Q1, 13
 , Q2, 12
 , Q3, 14
 , Q4, 11
 2012, Q1, 12
 , Q2, 14
 , Q3, 15
 , Q4, 17",
 show_col_types = FALSE
   )
)
```

```{r}
# Fill downwanrds (default)
tbl %>% fill(year)
```

```{r}
# See what happens when we fill upwards
tbl %>% fill(year, .direction = "up")
```

### Nesting Data

A tibble entry usually contains simple data, such as numbers or strings, but it 
can also hold complex data, such as other tables. Consider this table:

```{r}
(tbl<-
   tribble(
     ~A,~B,~C,
     1,11,21,
     1,11,12,
     2,42,22,
     2,15,22,
     2,15,32,
     4,13,32
   )
 )
```

From this table, we can create one that, for each value in the A column, contains a 
table of B and C for that A value.
```{r}
# Create a list-column of B and C
(nested_tbl <-
  tbl %>%
  nest(BC = c(B, C))
 )
```
You can do the reverse operation with `unnest()`
```{r}
# Unnest list-column
nested_tbl %>% unnest()
```

Nesting is not a typical operation in data analysis, but it is there if you need it.