---
title: "R4DS-Quickref: Ch4 -- Tidy Select"
output: html_notebook
---

So-called “tidy select” is not a package you would use on its own (although you can 
import it and it is called tidyselect), rather it is a small language for selecting 
columns in a data frame or tibble. We cannot, however, use the language without 
functionality from packages, so I will use the function `select` (from the `dplyr` package)
that helps you select columns from a tibble or a data frame.

Overview of selection features

Tidyverse selections implement a dialect of R where operators make it easy to select variables:

    : for selecting a range of consecutive variables.

    ! for taking the complement of a set of variables.

    & and | for selecting the intersection or the union of two sets of variables.

    c() for combining selections.

In addition, you can use selection helpers. Some helpers select specific columns:

    everything(): Matches all variables.

    last_col(): Select last variable, possibly with an offset.

These helpers select variables by matching patterns in their names:

    starts_with(): Starts with a prefix.

    ends_with(): Ends with a suffix.

    contains(): Contains a literal string.

    matches(): Matches a regular expression.

    num_range(): Matches a numerical range like x01, x02, x03.

These helpers select variables from a character vector:

    all_of(): Matches variable names in a character vector. All names must be present, otherwise an out-of-bounds error is thrown.

    any_of(): Same as all_of(), except that no error is thrown for names that don't exist.

This helper selects variables with a function:

    where(): Applies a function to all variables and selects those for which the function returns TRUE.


```{r}
library(dplyr)
```

```{r}
# Generate some data as a tibble
(tbl <- 
   tribble(
     ~sample,~min_size,~max_size,~min_weight,~max_weight,
     "foo",13,16,45.2,67.2,
     "bar",12,17,83.1,102.5
   )
 )
```

### Basic Select
```{r}
# Simply select columns by name
tbl %>% 
  select(sample, min_size, min_weight)
```

```{r}
# _You can also use 'c()' to combine wanted selection of columns
tbl %>% 
  select(c(sample, min_weight))
```


### Ranges

```{r}
# Select a range of contiguous columns using ':'
tbl %>% 
  select(min_size:max_weight)
```

```{r}
# Select multiple contiguous ranges using ':'
tbl %>% 
  select(sample:min_size, min_weight:max_weight)
```

### Complements

```{r}
# You can negate your selection using a bang '!':
tbl %>% 
  select(!sample)
```

```{r}
# Negating also works with ranges
# Parens are optional but it might be better to use then when negating to avoid surprises
tbl %>% 
  select(!(min_size:max_weight))
```

```{r}
tbl %>% 
  select(min_size:max_weight)
```

```{r}
# You can use '-' to negate too
tbl %>% 
  select(-sample)
```

```{r}
# To use '-' for a range you have to close it in parentheses
tbl %>% 
  select(-(sample:min_weight))
```

```{r}
# Note that this does not work with '!'
tbl %>% 
  select(min_size: max_weight, !min_size)
```

```{r}
# But this does work using '-' but the order changes
tbl %>% 
  select(min_size: max_weight, -min_size)
```

```{r}
# Select discontiguous columns using 'c()'
tbl %>% 
  select(c(min_size, max_weight))
```

```{r}
# Negate previous selection using '-'
tbl %>% 
  select(-c(min_size, max_weight))
```

```{r}
# Also wroks using '!'
tbl %>% 
  select(!c(min_size, max_weight))
```

### Unions and Intersections

```{r}
# It gets strange if you include the complements of columns and the columns themselves
# IMO this does not work as it should
tbl %>% 
  select(sample, !(sample:min_size))
```

```{r}
# I would prefer using '-' over '!'
tbl %>% 
  select(sample:max_weight, -(sample:min_size))
```

```{r}
# Explicitly ask for intersections using '&'
tbl %>% 
  select(sample:min_weight & max_size:max_weight)

```

```{r}
# Explicitly ask for unions using '|'
tbl %>% 
  select(sample:min_weight | max_size:max_weight)
```

```{r}
# Comma also produces an union
tbl %>% 
  select(sample:min_weight, max_size:max_weight)
```

### Select Columns Based on Name
```{r}
# starts_with() selects columns that start with the given string
tbl %>% 
  select(starts_with("min"))
```

```{r}
# Similarly, ends_with() can be used to select columns that end with the given string
tbl %>% 
  select(ends_with("weight"))
```

```{r}
# Use contains() to select strings that contain the given string
tbl %>% 
  select(contains("_s"))
```

```{r}
# The matches() function lets you select columns based on regular expressions
tbl %>% 
  select(matches(".*_.*"))
```

```{r}
# You can negate the selections with '!'
tbl %>% 
  select(!contains("_s"))
```

```{r}
# Or with '-'
tbl %>% 
  select(-contains("_s"))
```


### Everything
```{r}
# The function everything() select all columns. This can be useful in some cases
# Here we override a previous negative selection (note that sample is now the last column)
tbl %>% 
  select(-sample, everything())
```

### Indexing from the Last Column

```{r}
# To select the last column, use last_col()
tbl %>% 
  select(last_col())
```

```{r}
# But you can use last_col() to index from the right by giving an index
# Here 0 gives you the last column
tbl %>% 
  select(last_col(0))
```

```{r}
# But now you can index from right->left; however remember this is zero indexing now
# This gives the fifth column from left (first column from right)
tbl %>% 
  select(last_col(4)) 
```

```{r}
# You can also use last_col() in ranges
tbl %>% 
  select(last_col(3):last_col(2))
```

### Selecting from Strings
```{r}
# You can use the all_of() or any_of() functions to select with a vector of column names
vars <- c("min_size", "min_weight")
tbl %>% 
  select(all_of(vars))
```

The difference between the two functions is that all_of() considers it an error if 
vars contain a name that isn’t found in the table, while any_of() does not.

```{r}
# Any_of doesn't care if some of the columns are not there
tbl %>% 
  select(any_of(c(vars, "foo")))
```

```{r}
# All_of cares
tbl %>% 
  select(all_of(c(vars, "foo")))
```

### Selecting Columns Based on Their Content

Perhaps the most powerful selection function is where. You give it a function as an 
argument, that function is called with each column, and the columns for which the 
function returns TRUE are selected.
```{r}
# If you only want numeric columns
tbl %>% 
  select(where(is.numeric))
```

```{r}
# Select numeric columns where the maximum value is over 100
tbl %>% 
  select(where(\(x) is.numeric(x) && max(x) > 100))
```






















